Reanalysis Step 4: Propensity Score Analysis with Twang

### 2025-09-12 reanalysis-4-propensity_scores.qmd
balance stat

>>>>

## Average Largest Absolute Standardized Mean Difference

```{r avg-largest-smd}
# Extract the largest absolute SMD for each imputation
largest_smds <- numeric(n_imputations)

for (i in 1:n_imputations) {
  twang_model <- twang_results[[i]]$model

  # Get balance table which contains standardized mean differences
  tryCatch({
    bal_table <- bal.table(twang_model)

    # Extract the weighted (after matching) standardized effect sizes
    # The bal.table contains effect sizes for each covariate
    if (!is.null(bal_table$es.mean) && !is.null(bal_table$es.mean$max.es)) {
      largest_smds[i] <- bal_table$es.mean$max.es
    } else {
      # Fallback: use the summary method
      model_summary <- summary(twang_model)
      if (!is.null(model_summary) && "max.es" %in% names(model_summary)) {
        largest_smds[i] <- model_summary$max.es
      } else {
        largest_smds[i] <- NA
      }
    }
  }, error = function(e) {
    # If balance table fails, set to NA
    largest_smds[i] <- NA
  })
}

# Calculate average of largest SMDs across imputations
avg_largest_smd <- mean(largest_smds, na.rm = TRUE)
n_valid_smds <- sum(!is.na(largest_smds))

# Create summary table
smd_summary <- data.frame(
  Imputation = 1:n_imputations,
  Largest_Absolute_SMD = round(largest_smds, 4)
)

kable(smd_summary, 
      caption = "Largest absolute standardized mean difference by imputation",
      col.names = c("Imputation", "Largest |SMD|"))

# Display average
cat("\n**Average of largest absolute SMDs across imputations:**", round(avg_largest_smd, 4), "\n")
cat("**Number of valid SMD values:**", n_valid_smds, "out of", n_imputations, "\n")

if (avg_largest_smd < 0.1) {
  cat("**Interpretation:** Excellent balance (avg largest |SMD| < 0.1)\n")
} else if (avg_largest_smd < 0.25) {
  cat("**Interpretation:** Good balance (avg largest |SMD| < 0.25)\n") 
} else {
  cat("**Interpretation:** Poor balance (avg largest |SMD| ≥ 0.25) - consider alternative methods\n")
}
```
